Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-03-24 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Edits to introduction and beginning of results.

Most interfaces to gas-accepting sputter ion sources are a closed inlet design. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the sample volume and the length and diameter of the capillary. Flow may be adjusted by changing the pressure by varying the sample volume. We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere. This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary and is not adjustable. The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream. An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps. This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure. A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios. Salazar and Szidat, Roberts gasbench, other open split methods.

@@

@@

Source currents are variable by day and by target. "Good" currents for a given day vary between 5-10uA, and it's not yet clear what causes this variability. Alignment, source condition, and settings of the ionizer and cesium are all factors. Current also varies by target for a given set of gas and source conditions. Some targets are clearly "bad", producing currents much lower than other targets for a given set of conditions. Capillary clogging can be an issue, especially when sampling from vials with acid.

@@

Measurement of a small number of replicates shows that between sample reproducibility is around 1%. Samples displaced by helium have slightly higher variability, but more tests are needed to better assess reproducibility.

>>>>

Most interfaces to gas-accepting sputter ion sources are a closed inlet design. A mixture of sample and helium is introduced to the source through a capillary, and flow is controlled by the pressure in the inlet volume and the length and diameter of the capillary. Pressure is adjusted by changing either the inlet volume or the amount of gas in the inlet. We designed a open inlet where a capillary samples gas for delivery to the ion source in a flowing stream that exhausts to the atmosphere, often called an open split. This defines the delivery pressure as 1 ATM and therefore flow is controlled by the capillary parameters and is not easily adjustable during measurement. The flow of CO2 to the source may be varied by controlling the ratio of CO2 to helium in the sample stream. Changing the ratio of CO2 to helium changes the partial pressure of CO2 and therefore the carbon flow rate to the source.

An open inlet has the advantage of simplicity; flow of CO2 to the source is preset and there are no separate sample transfer and dilution steps. This may translate to greater throughput by eliminating the time required for sample transfer, dilution, and setting of delivery pressure. A disadvantage of the open split is that the flow of CO2 to the source varies with sample size (CO2:He) and changes over the course of a measurement due to dilution of the sample by displacement helium. Our method compensates for this by limiting CO2 flow to a range that produces relatively constant C- ion current and stable ratios. Salazar and Szidat, Roberts gasbench, other open split methods.

@@

Our initial work with the NEC source established optimal source parameters for consistent source performance. We used tank CO2 connected through a regulator to the ion source to measure source performance over a range of CO2 flows. Optimal flow for this source is 7.5uLs-1 CO2, but the source produces near optimal currents over a wide range of flows. Adding helium to the gas mixture produced similar currents to pure CO2 at the same CO2 flow rates.

@@

Source currents are variable by day and by target under similar source, inlet, and sample conditions. Capillary condition, source and gas delivery arm alignment, slight variability in target geometry, source condition, and performance of the ionizer and cesium may all be factors. Current also varies by target for a given set of gas and source conditions. Some targets are clearly "bad", producing currents much lower than other targets for a given set of conditions. Capillary clogging can be an issue, especially when sampling from vials with acid.

@@
